---------
title : "Reproducible Research: Peer graded assignment""
author: "Lalit Ojha"
date  : "January 19, 2018"
---------
```{r setup, include=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
```


__________________________________________________

Code for reading the dataset into R

```{r echo=TRUE}
activity.csv<-read.csv("activity.csv",sep=",") #reading data into R
activitydata<-subset(activity.csv, is.na(activity.csv$steps)==F) #reading data without NA values
```

Code for Calculating steps each day

```{r echo=TRUE}

totalstepsperday <- aggregate(steps ~ date, activitydata, sum)
```


Histogram of the total number of steps taken each day

```{r echo=TRUE}
hist(totalstepsperday$steps, main="Number of steps", xlab="Total no. of steps taken each day", ylab="No. of Days", col="grey")
```

Mean and median of steps taken each day

```{r echo=TRUE}
meansteps<-mean(totalstepsperday$steps)
mediansteps<-median(totalstepsperday$steps)
```

Time-series plot of the average number of steps taken by interval

```{r echo=TRUE}
avg <- aggregate(steps ~ interval, activitydata, mean)
plot(avg$interval, avg$steps, type='l', col='grey', main="Average number of steps by Interval", xlab="Time Intervals", ylab="Average number of steps")

```


The 5-Minute interval that, on average, contains the maximum number of steps:

```{r echo=TRUE}
intervalmax <- which.max(avg$steps)
print (paste("The interval with the highest avg steps is ", avg[intervalmax, ]$interval, " and the no of steps for that interval is ", round(avg[intervalmax, ]$steps, digits = 1)))

```


Calculating total rows containing missing values in original data

```{r echo=TRUE}

sum(is.na(activity.csv))

```

Data-Imputation:
Strategy: Using predictive mean matching(pmm)

```{r echo=TRUE}
install.packages(" mice")
library(mice)
md.pattern(activity.csv)
imputeactivity<-mice(activity.csv, m=5, maxit=5, method='pmm', seed=500) #imputing 5 datasets using pmm 
summary(imputeactivity)
mice(data = activity.csv, m = 5, method = "pmm", maxit = 5, seed = 500) #multiply ijmputed dataset
imputeactivity$imp$steps #checking imputed data
completeactivity <- complete(imputeactivity,2) #creating a complete dataset using second imputed dataset

```


Histogram of the total number of steps taken each day after missing values are imputed

```{r echo=TRUE}
totalstepsimputed <- aggregate(steps ~ date, completeactivity, sum)
hist(totalstepsimputed$steps, main="Total Number of steps (imputed)", xlab="Total no. of steps taken each day", ylab="No. of Days", col="grey")
```


Mean and Median steps per day in imputed data

```{r echo=TRUE}

mean(totalstepsimputed$steps)
mean(totalstepsperday$steps)

median(totalstepsimputed$steps)
median(totalstepsperday$steps)


```

Dividing the imputed data into weekdays and weekends
```{r echo=TRUE}
completeactivity$weekdays <- weekdays(as.Date(completeactivity$date))
completeactivity$weekdays <- ifelse(completeactivity$weekdays %in% c("Saturday", "Sunday"),"weekend", "weekday")

```

Calculating average steps of each interval
```{r echo=TRUE}
average<-aggregate(steps~interval+weekdays, completeactivity, mean)
```

Creating the plot
```{r echo=TRUE}
xyplot(steps ~ interval | weekdays, data = average, layout = c(1, 2), type="l", xlab = "Interval", ylab = "Number of steps", col='grey')

```


